---
title: "BellaBeat Analysis"
author: "Michael Nicolas"
date: "2023-05-30"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## GitHub Documents

This R Markdown document provides an analysis of the data collected from a fitness tracking application, FitBase. The dataset includes information on users' activity levels, sleep patterns, BMI, and other relevant metrics. The purpose of this document is to explore the relationship between sleep and activity levels, as well as tracking when users are mostly active during the day.


##  Code



```{r  path to the directory}
setwd("/cloud/project/")
```




```{r Loading the necessary libraries }
library(ggplot2)
library(dplyr)
library(tidyverse)
```

```{r sorting the data }

sleepData <- read.csv("sleepDay_updated.csv")

# Convert DayOfWeek column to factor with correct order
sleepData$DayOfWeek <- factor(sleepData$DayOfWeek, levels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))

# Create a new column with numeric representation of weekdays
sleepData$DayNum <- as.integer(sleepData$DayOfWeek)

# Aggregate the data by DayOfWeek and calculate the mean values
aggData <- aggregate(cbind(AverageMinutesAsleep = TotalMinutesAsleep, AverageTimeInBed = TotalTimeInBed, DayNum) ~ DayOfWeek, sleepData, FUN = mean)

# Order the aggregated data by the numeric representation of weekdays
orderedData <- aggData[order(aggData$DayNum), ]

# Remove the DayNum column from the final result
orderedData$DayNum <- NULL

# View the ordered data
orderedData


```
```{r Visualizing Average Minutes Asleep}
# Create a split bar plot for AverageMinutesAsleep (ordered descending)
minutes_asleep_plot <- ggplot(orderedData, aes(x = reorder(DayOfWeek, -AverageMinutesAsleep), y = AverageMinutesAsleep)) +
  geom_bar(stat = "identity", fill = "dark red", width = 0.6) +
  geom_text(aes(label = round(AverageMinutesAsleep)), vjust = -0.5) +
  labs(title = "Average Minutes Asleep by Day",
       x = "Day of the Week",
       y = "Average Minutes Asleep") +
  theme_minimal() +
  ylim(0, max(orderedData$AverageMinutesAsleep) * 1)

minutes_asleep_plot

```




```{r Visualizing Average Time in Bed}
# Create a bar plot with the number of minutes displayed above each bar
bar_plot <- ggplot(orderedData, aes(x = reorder(DayOfWeek,  -AverageTimeInBed), y=AverageTimeInBed) ) +
  geom_bar(stat = "identity", fill = "dark violet") +
  geom_text(aes(label = round(AverageTimeInBed)), vjust = -0.5) +
  labs(title = "Average Time Spent in Bed by Day of the Week",
       x = "Day of the Week",
       y = "Average Time in Bed (minutes)") +
  theme_minimal() +
  ylim(0, max(orderedData$AverageTimeInBed) * 1.1)

bar_plot

```
```{r}
TimeActive <- read_csv("UsersMostlyActive.csv", col_types = "c")



# Reorder the levels of the x-axis variable based on a clockwise order starting from 12:00 am
TimeActive$activity_time <- factor(TimeActive$activity_time,
                                   levels = c("12:00 AM", "1:00 AM", "2:00 AM", "3:00 AM", "4:00 AM", "5:00 AM",
                                              "6:00 AM", "7:00 AM", "8:00 AM", "9:00 AM", "10:00 AM", "11:00 AM",
                                              "12:00 PM", "1:00 PM", "2:00 PM", "3:00 PM", "4:00 PM", "5:00 PM",
                                              "6:00 PM", "7:00 PM", "8:00 PM", "9:00 PM", "10:00 PM", "11:00 PM"))

# Create the bar plot
barplot <- ggplot(TimeActive, aes(x = activity_time, y = average_intensity)) +
  geom_bar(stat = "identity", fill = "dark blue") +
  labs(title = "Users Most and Least Active Time of the Day",
       x = "Time",
       y = "Intensity") +
  theme_minimal() +
  geom_text(aes(label = activity_time, fontface = "bold"), vjust = -1.5, size = 1.8) +
  theme(axis.text.x = element_blank())

# Display the bar chart
print(barplot)

```







```{r comparison of Non-Active Minutes and Total Minutes Asleep}

 ActivityVsSleep <- read.csv("ActivityAndSleep.csv")

# Find the highest and lowest values of TotalMinutesAsleep
max_minutes_asleep <- max(ActivityVsSleep$TotalMinutesAsleep)
min_minutes_asleep <- min(ActivityVsSleep$TotalMinutesAsleep)

# Create the visualization with the custom line segment
ggplot(ActivityVsSleep, aes(x = non_ActiveMinutes, y = TotalMinutesAsleep)) +
  geom_point(color = "red") +
  geom_segment(aes(x = min(non_ActiveMinutes), y = max_minutes_asleep,
                   xend = max(non_ActiveMinutes), yend = min_minutes_asleep),
               color = "blue") +
  labs(x = "Non-Active Minutes", y = "Total Minutes Asleep") +
  ggtitle("comparison of Non-Active Minutes and Total Minutes Asleep
                             (Quality Sleep)") +
  theme_minimal() +
  xlim(0, 1500) +  # Adjust the x-axis range
  ylim(0, 1000)    # Adjust the y-axis range


```



```{r reading table}
activity_vs_bmi <- read.csv("Activity_BMI.csv")
```

```{r comparing ActiveMinutes vs avg_BMI}

# Calculate the sum of minutes spent active
activity_vs_bmi$ActiveMinutes <- activity_vs_bmi$avg_VeryActiveMinutes +
                                activity_vs_bmi$avg_FairlyActiveMinutes

# Calculate the average non-active minutes
activity_vs_bmi$avg_nonactiveminutes <- activity_vs_bmi$avg_SedentaryMinutes + activity_vs_bmi$avg_LightlyActiveMinutes

# Create a line plot comparing avg_nonactiveminutes vs avg_BMI
plot_nonactive <- ggplot(activity_vs_bmi, aes(x =  avg_nonactiveminutes, y = avg_BMI)) + 
  
  geom_point(color = rainbow(nrow(activity_vs_bmi)), size = 4)+
  labs(y = "Average BMI", x = "Average Non-Active Minutes") 
  
  

# Create a line plot comparing ActiveMinutes vs avg_BMI
plot_active <- ggplot(activity_vs_bmi, aes(x = ActiveMinutes, y = avg_BMI )) +
  geom_point(color = rainbow(nrow(activity_vs_bmi)), size = 4) +
  labs(y = "Average BMI", x = "Minutes Spent Active") 
  

# Display the plots side by side
library(gridExtra)
grid.arrange(plot_nonactive, plot_active, ncol = 2 , top = "Comparison of Activity and Non-Activity by BMI")



```




```{r  relationship between activity level and calories burnt}
library(ggplot2)
library(cowplot)

# Create subset data based on the condition
subset_data <- subset(activity_data, (VeryActiveMinutes > 0 | FairlyActiveMinutes > 0 | LightlyActiveMinutes > 0 | SedentaryMinutes > 0) & Calories > 0)

# Create four separate visuals comparing each column with calories
plot1 <- ggplot(subset_data, aes(x = VeryActiveMinutes, y = Calories)) +
  geom_point(color = "dark cyan") +
  geom_smooth(method = "lm", se = FALSE, color = "red") 

plot2 <- ggplot(subset_data, aes(x = FairlyActiveMinutes, y = Calories)) +
  geom_point(color = "dark cyan") +
  geom_smooth(method = "lm", se = FALSE, color = "orange") 

plot3 <- ggplot(subset_data, aes(x = LightlyActiveMinutes, y = Calories)) +
  geom_point(color = "dark cyan") +
  geom_smooth(method = "lm", se = FALSE, color = "yellow") 

plot4 <- ggplot(subset_data, aes(x = SedentaryMinutes, y = Calories)) +
  geom_point(color = "dark cyan") +
  geom_smooth(method = "lm", se = FALSE, color = "green") 

# Arrange the four plots in a single row
combined_plot <- plot_grid(plot1, plot2, plot3, plot4, nrow = 1)

# Display the combined plot
print(combined_plot)

```


